---
title: "Assignment 1"
author: "Jiri Petr"
date: "Thursday, April 16, 2015"
output: html_document
---


### Loading and processing the data

```{r Loading Data}
setwd("C:/users/jpet/coursera/datascience/reproducibleresearch")
data <- read.csv("activity.csv")
```

```{r Cleaning data}
dataclear <- data[complete.cases(data),]
```

### What is mean total number of steps taken per day?


```{r}
library(dplyr)
hgram <- summarize(group_by(dataclear, date), sum(steps))
colnames(hgram) <- c("day", "steps")
hist(hgram$steps, 15)
mean <- mean(hgram$steps)
median <- median(hgram$steps)
```

The daily mean of steps is `r mean` and median of steps in one day is `r median`.


 ### What is the average daily activity pattern?
 
 
```{r}
sminutes <- summarize(group_by(dataclear, interval), mean(steps))
colnames(sminutes) <- c("interval", "steps")
plot(sminutes$interval, sminutes$steps, type = "l")
max <- sminutes[sminutes$steps == max(sminutes$steps),1]
```

 On average, maximum steps are taken in five minute interval with beginning at `r max`.
 
 
 ### Imputing missing values
 
 
```{r}
nas <- sum(!complete.cases(data))
```

Total number of incomplete rows is `r nas`.

All missing values will be filled by median of particular interval. Let's look what happen?

```{r Imputing missing values}
for (i in 1:length(data[,1])) {
        if (is.na(data[i,1]) == TRUE) {
                data[i,1] <- sminutes[sminutes$interval == data[i,3],2]
        }
}

datanew <- summarize(group_by(data, date), sum(steps))
colnames(datanew) <- c("day", "steps")
mean2 <- mean(datanew$steps)
median2 <- median(datanew$steps)
```

Previous means and median was `r mean` and `r median` respecitively. After inserting estimated valuest instead of NAs, mean and median of steps per day is `r mean2`, `r median2` respectively.


### Are there differences in activity paterns between weekdays and weekends?


``` {r weekdays vs. weekends}
for (i in 1:length(data[,1])) {
        if (strftime(data[i,2], '%u') < 6) {
                data[i,4] <- "weekday"
        } else {
                data[i,4] <- "weekend"
        }
}

colnames(data)[4] <- "wday"
weekends <- summarize(group_by(data,interval,wday), sum(steps))
colnames(weekends) <- c("interval","wday", "steps")

library(ggplot2)

qplot(interval, steps,data = weekends, facets = ~ wday, geom = "line")

